---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pdftabler

<!-- badges: start -->
<!-- badges: end -->

The goal of pdftabler is to 
create arranged data frame extracted from PDF file by R.
It deals especially with time-series data.
This package is required pdftools.

## Installation

You can install the development version of pdftabler from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("MTRBEKTCBT/pdftabler")
```

## Example

```{r}
library(pdftabler)
```

The data extracted from a package `pdftools::pdf_text` is 
often dirty by unnecessary spaces 
and is a vector rather than a dataframe.
This is a basic example which shows you how to 
get the data obtained from pdf to a complete dataframe.

Suppose we already have vector data extracted by `pdftools::pdf_text` that has just been extracted.
We'll use `conbini` data which records 
[JFA sales from 2005 to 2021](https://www.jfa-fc.or.jp/particle/320.html)
to show how each function works.

First of all, the data has to be split into each cell of table
which originally displayed in PDF.
To process this task, use `pt_cell()` and `pt_split()`.

```{r}
conbini |>
  pt_cell() |>
  purrr::flatten_chr() |>
  pt_split(initial = 29, n_col = 15, dur_month = 2)
```

The argument `initial` in `pt_split()`
denotes a initial cell for records.
Note that it indicates a record for inner data,
do not specify a raw of column names.

We will also wonder which number is to be `initial`?
We can use `pt_index` to resolve this task.
It creates index dataframe that indicates 
which place any word or symbol is in the vector.

```{r}
key <- sprintf("%d月", 1:12)

idx <- 
  conbini |>
  pt_cell() |>
  purrr::flatten_chr() |>
  pt_index(keywords = key)

idx |> head()
```

Now we can use another function `pt_split_multi` that can handle dataframe as argument.
It is a recursive version of function `pt_split`.


```{r}
df <- 
conbini |>
  pt_cell() |>
  purrr::flatten_chr() |>
  pt_split_multi(index = idx)

df |> head()
```

Then, we may also complain about data type, 
that all of columns in data are character but
we can easily modify this to use some methods already exist.
Since the example data is essentially numerical and 
limited in time period,
I packed a method for it that processes required task at one step.

```{r}
df |> pt_modify() |> head()
```

Now we can analyze the data as always.

```{r}
library(ggplot2)
df |> 
  pt_modify() |> 
  dplyr::mutate(date = stringr::str_c(year, month, "01") |> lubridate::ymd()) |> 
  ggplot(aes(date, V5)) +
  geom_line() +
  labs(
    title = "コンビニエンスストアの売上高前年比の推移",
    x = "月",
    y = "既存店売上高前年比"
  )
```

